Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-10 mr-review.Rmd
added sensitivity analysis text

@burgess_sensitivity_2017 divided the suite of methods for sensitivity analysis 
into two categories: 1) those that support the validity of causal inferences in a MR 
study and 2) those that question that validity. Because not all confounding 
variables are known or measured, assumptions 2 and 3 are not fully verifiable 
in MR studies. Recognition of this fact has inspired the development of 
sensitivity analysis tools. While we can't assess whether a SNP instrument is 
associated with every confounder of the exposure-outcome association, it is 
possible to examine the associations of the SNP instrument with the measured 
covariates. While absence of such associations doesn't guarantee satisfaction 
of the assumption, presence of SNP-covariate associations must be investigated
carefully, as they may constitute assumption violations [@burgess_sensitivity_2017].

>>>>

Because not all confounding 
variables are known or measured, assumptions 2 and 3 are not fully verifiable 
in MR studies. Recognition of this fact has inspired the development of 
sensitivity analysis tools. While we can't assess whether a SNP instrument is 
associated with every confounder of the exposure-outcome association, it is 
possible to examine the associations of the SNP instrument with the measured 
covariates. While absence of such associations doesn't guarantee satisfaction 
of the assumption, presence of SNP-covariate associations must be investigated
carefully, as they may constitute assumption violations [@burgess_sensitivity_2017].

@burgess_sensitivity_2017 considered a collection of sensitivity analysis methods when 
working with GWAS summary statistics. @burgess_sensitivity_2017 presented methods for both assessing the
MR assumptions, to the extent possible, and performing robust analyses. For example, 
they used measured covariates to assess for possible associations with the SNP 
instruments. While they can't rule out the possibility of unmeasured confounding, they can study the possibility of measured covariates serving as confounders.

To illustrate their approach, @burgess_sensitivity_2017 shared a case study in which they examined the causal effect of 
C-reactive protein (CRP) levels on coronary artery disease risk with four genetic 
variants in the CRP gene region and 17 other genetic variants that affect coronary 
artery disease risk. They use measured covariates to probe for SNP-covariate 
associations, and followed it up with scatter plots and Cochran's Q test on the causal 
estimates to inquire about whether the 
SNP instruments all identify the same causal parameter [@greco_m_detecting_2015;@small_sensitivity_2007;@higgins_measuring_2003]. 
Additionally, @burgess_sensitivity_2017 suggests using a funnel plot, like those in the 
meta-analysis literature, to visualize possible evidence of directional pleiotropy, where the average pleiotropic effects of the SNP instruments is nonzero [@sterne_recommendations_2011]. Additionally, Egger regression can be useful in this 
setting [@egger_bias_1997;@bowden_mendelian_2015].
